---
output:
  pdf_document: default
  html_document: default
---
This tutorial take CITE-seq for example. We load the simulation results saved in "../output/simulation_result/" folder as input. The output is the umap of real data, umap of simulated data, and umap of  

load package
```{r}
library(SingleCellExperiment)
library(data.table)
library(scater)
library(glue)
library(scran)
library(ggplot2)
library(ggpubr)
library(umap)
make_sce <- function(expr, label){
  sce <- SingleCellExperiment(list(logcounts=expr))
  sce$celltype <- as.factor(label)
  return(sce)
}
```

```{r}
dataset_name = "CITEseq"
```

load real and simulated data
```{r}
nfeatures_rna = 10249
nfeatures_adt = 189
feature_num = nfeatures_rna + nfeatures_adt
      
train.sim <- t(fread(glue("../output/simulation_result/CITEseq/sim.csv")))
train.sim <- train.sim[2:dim(train.sim)[1],]
train.rna.sim <- train.sim[1:nfeatures_rna,]
train.adt.sim <- train.sim[(nfeatures_rna+1): feature_num,]
colnames(train.rna.sim)<-glue("sim_{c(1:dim(train.rna.sim)[2])}")
colnames(train.adt.sim)<-glue("sim_{c(1:dim(train.adt.sim)[2])}")

train.real <- t(fread(glue("../output/simulation_result/CITEseq/real.csv")))
train.real <- train.real[2:dim(train.real)[1],]
train.rna.real <- train.real[1:nfeatures_rna,]
train.adt.real <- train.real[(nfeatures_rna+1): feature_num,]
colnames(train.rna.real)<-glue("sim_{c(1:dim(train.rna.real)[2])}")
colnames(train.adt.real)<-glue("sim_{c(1:dim(train.adt.real)[2])}")
    
#train.cty.sim <-  t(fread(glue("../output/simulation_result/CITEseq/label.csv")))[2,]  #label information
```

Then, do umap for real and simulated RNA data
```{r}
rna.combined <- cbind(train.rna.real, train.rna.sim)
sce.rna.combined <- make_sce(rna.combined, c(rep("original",dim(train.rna.real)[2]),rep("augment",dim(train.rna.sim)[2])))
sce.rna.combined <- runUMAP(sce.rna.combined, n_threads=10)
umap_both_rna_layout <- sce.rna.combined@int_colData$reducedDims$UMAP
umap_both_rna_layout <- as.data.frame(umap_both_rna_layout)
data <- c(rep("2",dim(train.rna.real)[2]),rep("4",dim(train.rna.sim)[2]))
umap_both_rna_layout1 <- umap_both_rna_layout[1:dim(train.rna.real)[2],]
umap_both_rna_layout2 <- umap_both_rna_layout[(dim(train.rna.real)[2]+1):(2*dim(train.rna.real)[2]),]
```

Next, show 1) umap of overlap real and simulated RNA data; 2) umap for real RNA data; 3) umap of simulated RNA data
```{r}
plot0<-ggplot(umap_both_rna_layout,aes(V1,V2,color=data)) +
  geom_point(size = 0.8,alpha=0.05,show.legend = FALSE)+ labs(x = "", y = "", title = "") + theme_classic()  + 
  theme(aspect.ratio = 1)+ theme(panel.grid =element_blank()) +theme(axis.text = element_blank()) +  theme(axis.ticks = element_blank()) +  theme(plot.title = element_text(hjust = 0.5,size=8),panel.spacing = unit(0, "lines")) + ggtitle("Real+simulate")
plot1 <- ggplot(umap_both_rna_layout1,aes(V1,V2)) +
  geom_point(size = 0.8,colour=2)+ labs(x = "", y = "", title = "") + theme_classic()  + 
  theme(aspect.ratio = 1)+ theme(panel.grid =element_blank()) +theme(axis.text = element_blank()) +  theme(axis.ticks = element_blank()) +  theme(plot.title = element_text(hjust = 0.5,size=8),panel.spacing = unit(0, "lines")) + ggtitle("Real RNA")
plot2 <- ggplot(umap_both_rna_layout2,aes(V1,V2)) +
  geom_point(size = 0.8,colour=4)+ labs(x = "", y = "", title = "") + theme_classic()  + 
  theme(aspect.ratio = 1)+ theme(panel.grid =element_blank()) +theme(axis.text = element_blank()) +  theme(axis.ticks = element_blank()) +  theme(plot.title = element_text(hjust = 0.5,size=8),panel.spacing = unit(0, "lines")) + ggtitle("simulate RNA")
top1<-ggarrange(plot0,plot1,plot2,ncol=3,nrow=1)
top1
```

Next, show 1) umap of overlap real and simulated ADT data; 2) umap for real ADT data; 3) umap of simulated ADT data
```{r}
adt.combined <- cbind(train.adt.real, train.adt.sim)
sce.adt.combined <- make_sce(adt.combined, c(rep("original",dim(train.adt.real)[2]),rep("augment",dim(train.adt.sim)[2])))
sce.adt.combined <- runUMAP(sce.adt.combined, n_threads=10)
umap_both_adt_layout <- sce.adt.combined@int_colData$reducedDims$UMAP
umap_both_adt_layout <- as.data.frame(umap_both_adt_layout)
data <- c(rep("2",dim(train.adt.real)[2]),rep("4",dim(train.adt.sim)[2]))
umap_both_adt_layout1 <- umap_both_adt_layout[1:dim(train.adt.real)[2],]
umap_both_adt_layout2 <- umap_both_adt_layout[(dim(train.adt.real)[2]+1):(2*dim(train.adt.real)[2]),]

plot4<-ggplot(umap_both_adt_layout,aes(V1,V2,color=data)) +
  geom_point(size = 0.8,alpha=0.05,show.legend = FALSE)+ labs(x = "", y = "", title = "") + theme_classic()  + 
  theme(aspect.ratio = 1)+ theme(panel.grid =element_blank()) +theme(axis.text = element_blank()) +  theme(axis.ticks = element_blank()) +  theme(plot.title = element_text(hjust = 0.5,size=8),panel.spacing = unit(0, "lines")) + ggtitle("Real+simulate")
plot5 <- ggplot(umap_both_adt_layout1,aes(V1,V2)) +
  geom_point(size = 0.8,colour=2)+ labs(x = "", y = "", title = "") + theme_classic()  + 
  theme(aspect.ratio = 1)+ theme(panel.grid =element_blank()) +theme(axis.text = element_blank()) +  theme(axis.ticks = element_blank()) +  theme(plot.title = element_text(hjust = 0.5,size=8),panel.spacing = unit(0, "lines")) + ggtitle("Real adt")
plot6 <- ggplot(umap_both_adt_layout2,aes(V1,V2)) +
  geom_point(size = 0.8,colour=4)+ labs(x = "", y = "", title = "") + theme_classic()  + 
  theme(aspect.ratio = 1)+ theme(panel.grid =element_blank()) +theme(axis.text = element_blank()) +  theme(axis.ticks = element_blank()) +  theme(plot.title = element_text(hjust = 0.5,size=8),panel.spacing = unit(0, "lines")) + ggtitle("simulate adt")
top2<-ggarrange(plot4,plot5,plot6,ncol=3,nrow=1)
top2  
```

